---
title: "Promise Zone Building Permits"
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#ffffff"
      fg: "#5E2C25"
      primary: "#444444"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
library(rphl)
library(mapview)
library(plotly)
library(ggthemr)
library(tmap)
library('downloadthis')
library(RColorBrewer)
library(scales)

tmap_options(basemaps = "Esri.WorldGrayCanvas")

ggthemr("pale") #set global ggplot theme

#"#444444" "#de6757" "#EB9050" "#3262AB" "#FF8D7D" "#C8E370" "#C45B4D" "#41a65c" "#5E2C25" "#78695F"
```

```{r data imports, cache = TRUE, include = FALSE}
base_url = "https://phl.carto.com/api/v2/sql"

query = "
        select *
        from permits
        where permitissuedate >= '2022-01-01'
        "
  
permits = st_as_sf(get_carto(query,
                              format = 'csv',
                              base_url = base_url,
                              stringsAsFactors = FALSE)|>
                      filter(!is.na(geocode_x),
                             !is.na(geocode_y)),
                      coords = c('geocode_x',
                                 'geocode_y'),
                      crs = st_crs('EPSG:2272'))

pz = st_read("C:/Users/Nissim/Documents/Building Permits for Dante/west_phl_building_permits/west_phl_building_permits/Philadelphia_Promise_Zone.shp") |>
        st_transform(crs = st_crs("EPSG:2272"))

pz_permits = permits[pz, ]

pz_construction_permits = pz_permits |>
                            filter(permittype %in% c("BUILDING", "DEMOLITION", "RESIDENTIAL BUILDING"))

pz_construction_permits$week = lubridate::week(pz_construction_permits$permitissuedate)

pz_construction_permits$permitissuedate = as.Date(pz_construction_permits$permitissuedate)

pz_construction_permits$status[pz_construction_permits$status == "AMENDMENT READY FOR ISSUE"] = "AMEND. READY FOR ISSUE"

pz_construction_permits$status[pz_construction_permits$status == "AMENDMENT REVIEW"] = "AMEND. REVIEW"
```

Column {data-width=200}
-------------------------------------

### Content

##### About the Dashboard
This dashboard displays building permits issued in the West Philadelphia in the past year. Data were last updated on `r Sys.Date()`. The maps and graphs are interactive and can be filtered by clicking on the relevant areas.

##### Download Data
To download a list of building permits in the Promise Zone in the last year, click the link below:
\
\
\
```{r download}
download_this(pz_construction_permits,
    output_name = "Promise Zone Building Permits",
    output_extension = ".xlsx",
    button_label = "Download Promise Zone Building Permits",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save")
```
\
\
Dashboard built by [Nissim Lebovits](nissim@desiign.upenn.edu).

Column {data-width=600}
-----------------------------------------------------------------------

### **Building Permits Mapped**

```{r map}
tmap_mode('view')

map_pal = c("#3262AB", "#FF8D7D","#41a65c")

tm_shape(pz_construction_permits) +
  tm_dots(title = "Permit Type",
          col = 'permittype',
          palette = map_pal,
          alpha = 0.5,
          size = 0.1,
          id = "address",
              popup.vars = c("Owner" = "opa_owner",
                             "Contractor" = "contractorname",
                             "Date Issued" = "permitissuedate",
                             "Permit Type" = "permittype",
                             "Permit Description" = "permitdescription",
                             "Type of Work" = "typeofwork",
                             "Status" = "status",
                             "Applicant Type" = "applicanttype")) +
tm_shape(pz) +
  tm_borders() +
tm_view(set.view = 15,
        view.legend.position = c("right", "bottom"))
```


Column {data-width=400}
-----------------------------------------------------------------------

### **Building Permits over Time**
*Hover over the plot to interact with it. Click the camera icon to download it as a .png file.*
```{r density}
ggplotly(
  ggplot(pz_construction_permits, aes(x = permitissuedate, fill = permittype)) +
  geom_density(alpha = 0.3, color = NA)+
    labs(x = "Date",
         y = "Count",
         fill = "Permit Type") +
  theme(legend.position = 'bottom') +
  scale_y_continuous(labels = percent) +
  scale_fill_manual(values = map_pal)
    )
```

### **Building Permits by Type of Work**
*Hover over the plot to interact with it. Click the camera icon to download it as a .png file.*
```{r}
ggplotly(
  ggplot(pz_construction_permits) +
    geom_bar(aes(x = typeofwork, fill = permittype), alpha = 0.4, position = 'stack') +
    labs(y = "Count",
         fill = "Permit Type") +
    scale_fill_manual(values = map_pal) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 7))
)
```

### **Building Permits by Status**
*Hover over the plot to interact with it. Click the camera icon to download it as a .png file.*
```{r}
ggplotly(
  ggplot(pz_construction_permits) +
    geom_bar(aes(x = status), fill = "#41a65c", alpha = 0.4) +
    labs(y = "Count") +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(angle = 45, hjust = 1, size = 7))
)
```