---
title: "Promise Zone Building Permits"
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#ffffff"
      fg: "#444444"
      primary: "#444444"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

```{=html}
<style type="text/css">
.value-box .value {
    color: white;
}
.value-box .caption {
    color: white;
}
</style>
```

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(sf)
library(rphl)
library(plotly)
library(ggthemr)
library(tmap)
library(downloadthis)
library(RColorBrewer)
library(scales)
library(tigris)
library(monochromeR)
library(verbaliseR)

tmap_options(basemaps = "Esri.WorldGrayCanvas")

ggthemr("pale") #set global ggplot theme

blue = generate_palette("#3262AB", modification = "go_lighter", n_colours = 4, view_palette = T)

green = generate_palette("#41a65c", modification = "go_lighter", n_colours = 4, view_palette = T)

red = generate_palette("#FF8D7D", modification = "go_lighter", n_colours = 4, view_palette = T)

map_pal_plus = paste(c(blue[1], blue[3], green[1], green[3], red[1], red[3]))

#"#444444" "#de6757" "#EB9050" "#3262AB" "#FF8D7D" "#C8E370" "#C45B4D" "#41a65c" "#5E2C25" "#78695F"

```

```{r data imports, cache = TRUE, include = FALSE}
base_url = "https://phl.carto.com/api/v2/sql"

query = "
        select *
        from permits
        where permitissuedate >= '2022-01-01'
        "
  
permits = st_as_sf(get_carto(query,
                              format = 'csv',
                              base_url = base_url,
                              stringsAsFactors = FALSE)|>
                      filter(!is.na(geocode_x),
                             !is.na(geocode_y)),
                      coords = c('geocode_x',
                                 'geocode_y'),
                      crs = st_crs('EPSG:2272'))

pz = st_read("C:/Users/Nissim/Documents/Building Permits for Dante/west_phl_building_permits/west_phl_building_permits/Philadelphia_Promise_Zone.shp") |>
        st_transform(crs = st_crs("EPSG:2272"))

pz_permits = permits[pz, ]

pz_construction_permits = pz_permits |>
                            filter(permittype %in% c("BUILDING", "DEMOLITION", "RESIDENTIAL BUILDING"))

pz_construction_permits$week = lubridate::week(pz_construction_permits$permitissuedate)

pz_construction_permits$permitissuedate = as.Date(pz_construction_permits$permitissuedate)

pz_construction_permits$permittype[pz_construction_permits$permittype == "BUILDING"] = "COMMERCIAL"

pz_construction_permits$permittype[pz_construction_permits$permittype == "RESIDENTIAL BUILDING"] = "RESIDENTIAL"

pz_construction_permits$status[pz_construction_permits$status == "AMENDMENT READY FOR ISSUE"] = "AMEND. READY FOR ISSUE"

pz_construction_permits$status[pz_construction_permits$status == "AMENDMENT REVIEW"] = "AMEND. REVIEW"

pz_construction_permits$status[pz_construction_permits$status == "AMENDMENT APPLICATION INCOMPLETE"] = "AMEND. APP. INCOMPLETE"

pz_construction_permits = pz_construction_permits |>
                            mutate(maj_or_min = case_when(
                                                          permittype == "COMMERCIAL" & typeofwork == "ADDITION AND/OR ALTERATION" ~ "Minor Commercial",
                                                          permittype == "RESIDENTIAL" & typeofwork == "ADDITION AND/OR ALTERATION" ~ "Minor Residential",
                                                          permittype == "COMMERCIAL" & typeofwork == "NEW CONSTRUCTION" ~ "Major Commercial",
                                                          permittype == "RESIDENTIAL" & typeofwork == "NEW CONSTRUCTION" ~ "Major Residential",
                                                          typeofwork == "MINOR DEMOLITION" ~ "Minor Demolition",
                                                          typeofwork == "MAJOR DEMOLITION" ~ "Major Demolition"
                                                        ))

pz_construction_permits$maj_or_min = factor(pz_construction_permits$maj_or_min, levels = c("Major Commercial",
                                                                                            "Minor Commercial",
                                                                                            "Major Residential",
                                                                                            "Minor Residential",
                                                                                            "Major Demolition",
                                                                                            "Minor Demolition"))
```
  
  
## Column {data-width="250"}

### About

###### **What am I looking at?**
This dashboard displays building permits issued in the West Philadelphia Promise Zone since `r prettify_date((lubridate::ymd(Sys.Date()) - lubridate::years(1)), uk_or_us = "US")`. Click on the maps and graph to interact with them.

###### **Download Data**
To download a list of building permits in the Promise Zone in the last year, click the link below:
\
\
\
```{r download}
download_this(pz_construction_permits,
    output_name = "Promise Zone Building Permits",
    output_extension = ".xlsx",
    button_label = "Download Promise Zone Building Permits",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save")
```
\
\

###### **Credits**

Dashboard built by [Nissim Lebovits](https://twitter.com/NissimLebovits). For questions, feedback, or requests, contact me at [nissim\@design.upenn.edu](mailto:nissim@design.upenn.edu){.email}.

### Active Commercial Building Permits

```{r building tally}
valueBox(format(nrow(pz_construction_permits |> filter(permittype == "COMMERCIAL")), big.mark = ","), icon = "fa-building", color = "#3262AB")
```
### Active Residential Building Permits

```{r residential building tally}
valueBox(format(nrow(pz_construction_permits |> filter(permittype == "RESIDENTIAL")), big.mark = ","), icon = "fa-home", color = "#41a65c")
```

### Active Demolition Permits

```{r demolition tally}
valueBox(format(nrow(pz_construction_permits |> filter(permittype == "DEMOLITION")), big.mark = ","), icon = "fa-bomb", color = "#FF8D7D")
```

## Column {data-width="450"}

### **Building Permits Mapped**

```{r map}
tmap_mode('view')

tm_shape(pz_construction_permits) +
  tm_dots(title = "Type of Work",
          col = 'maj_or_min',
          palette = map_pal_plus,
          alpha = 0.7,
          size = 0.1,
          id = "address",
              popup.vars = c("Owner" = "opa_owner",
                             "Contractor" = "contractorname",
                             "Date Issued" = "permitissuedate",
                             "Permit Type" = "permittype",
                             "Permit Description" = "permitdescription",
                             "Type of Work" = "typeofwork",
                             "Status" = "status",
                             "Applicant Type" = "applicanttype")) +
tm_shape(pz) +
  tm_borders() +
tm_view(set.view = 15,
        view.legend.position = c("right", "bottom"))
```

## Column {data-width="300"}

### Building Permits Issued per Week
```{r typeofwork hist}
ggplotly(
  ggplot(pz_construction_permits, aes(x = permitissuedate, fill = maj_or_min)) +
  geom_histogram(alpha = 0.7, color = NA, bins = 52) +
  labs(y = "Permits Issued") +
  facet_grid(maj_or_min ~ .) + #arranges charts in rows instead of side by side
  theme(legend.position = 'none',
        axis.title.x = element_blank())+
  scale_fill_manual(values = map_pal_plus)
)
```